---
title: "Assignment 3"
output: html_notebook
---

Libraries
```{r}
library(plm)
library(ggplot2)
```


Import Data
```{r}
data <- read.csv("~/Desktop/Assignments - Empirical Real Estate/Assignment 3/Bulwien_Gesa.txt", header = T, sep = "")
```


==========
Exercise 1
==========


Create laged rent
```{r}
data <- pdata.frame(data, index=c("City", "Jahr"))

data$rent_lag <- lag(data$rent, 1)
data$stock_lag <- lag(data$stock, 1)

```

Run panel regression to test the presence of fixed effects
```{r}

fit_ols <- lm(rent ~ rent_lag + stock_lag + stock + v, data = data)

summary(fit_ols)


fit_panel <- plm(rent ~ rent_lag + stock_lag + stock + v, data = data, model = "within")

summary(fit_panel)

fixef(fit_panel, type = "level")
```


Create Rent factor variable for different rent contract types
```{r}

panel_rent_type <- plm(log(rent) ~ log(rent_lag) + log(stock_lag) + log(stock) + log(v) + rent_lag:City, data = data, model = "within")

summary(panel_rent_type)

```

Controlling for both times and city fixed effects
```{r}
panel_rent_times <- plm(log(rent) ~ log(rent_lag) + log(stock_lag) + log(stock) + log(v) + rent_lag:City, data = data, effect = "twoways")

summary(panel_rent_times)

```


==========
Exercise 2
==========

Select the subset Frankfurt, Hamburg, Berlin, Stuttgart
```{r}
data_sub <- data[data$City %in% c("Frankfurt.(Main)", "Hamburg", "Berlin", "Stuttgart"), ]

data_sub$City <- factor(data_sub$City) #so that the City variables get the new levels argument
```

Regression on lagged vacancy
```{r}

panel_sub <- plm(log(v) ~ log(lag(v,1)), data = data_sub, effect = "twoways")

summary(panel_sub)

fixef(panel_sub)

resi <- residuals(panel_sub)

data_sub <- data_sub[-which(data_sub$Jahr=="1990"),]

data_sub$fit <- data_sub$v - resi 
data_sub$residuals <- resi

```

Plot Vacancy rate vs. Fitted value
```{r}
ggplot(data_sub, aes(x = Jahr)) + geom_point(aes(y = fit), color ="darkblue", alpha = 0.3) + geom_point(aes(y = v), color ="red", alpha =0.3) +  facet_wrap(~ City) +  scale_x_discrete("Cut", labels = abbreviate)+ labs(title = "Realization vs. Prediction", x ="Year" , y="Realization and Prediction") + theme_bw() + scale_x_discrete(breaks = c("1990", "1995", "2000", "2005"))
```

Using Normal plots
```{r}
par(mfrow= c(2,2))
for (i in levels(data_sub$City)){
  
  dat <- data_sub[data_sub$City %in% i, c("v", "Jahr", "fit")]
  
  plot(dat$v ~ dat$Jahr, type = "p", color = "red", ylab = "Realized vs. Predicted", xlab ="Jahr")
  par(new=T)
  plot(dat$fit ~ dat$Jahr, type = "p", color = "blue", axes = F, ylab = "", xlab = "")
  title(i)
}
```



Use the tapply() function to compute the root mean squared error (RMSE) for each city. Remember that you can program a function in tapply "on the fly" using tapply(numeric variable, categorical variable, function(x) .).

```{r}
RMSE_fun <- function(x){
  rmse <- sqrt(sum(x**2)/length(x))
}

RMSE <- tapply(data_sub$residuals, data_sub$City, RMSE_fun)
RMSE
```



